---
title: "Eukdetect output analysis"
format: 
  html:
    author: "Johnathan Shih"
    date: today
    code-fold: true
    page-layout: full
    toc: true
    toc-location: "left"
    smooth-scroll: true
execute:
  cache: false
---


```{r}
#| warning: false
#| message: false
#| fig-width: 6
#| fig-height: 6

library(purrr)
library(cowplot)
library(tidyverse)

source("eukdetect_analysis_functions.R")

cohorts <- c("predict1","predict2","predict3_uk",
             "predict3_us1","predict3us22a","hadza")

# All intensive analyses were run with a cluster compute node with eukdetect_analysis_pace.R

print(cohorts)



p_val_threshold <- 0.01
cohort_inclusion <- 2
```


Following figures use adjustable filters of p`.adj.hochberg > 0.01` in at least 2 cohorts.

### Difference in clr-transformed bacterial abundance ~ Blastocystis

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 15

p_blast <- read_csv("data/predict1_blast_interactions.csv") %>% 
  mutate(cohort = "predict1") %>%
  bind_rows(
    read_csv("data/hadza_blast_interactions.csv") %>% 
    mutate(cohort = "hadza"),
    read_csv("data/predict3_uk_blast_interactions.csv") %>% 
    mutate(cohort = "predict3_uk"),
    read_csv("data/predict2_blast_interactions.csv") %>% 
    mutate(cohort = "predict2"),
    read_csv("data/predict3_us1_blast_interactions.csv") %>% 
    mutate(cohort = "predict3_us1"),
    read_csv("data/predict3us22a_blast_interactions.csv") %>% 
    mutate(cohort = "predict3us22a")
  ) %>%
  filter(p.adj.hochberg < p_val_threshold) %>% # only significant taxa
  group_by(taxa) %>%
  filter(n() > cohort_inclusion) %>% # only plotting significant taxa in at least n cohorts
  ungroup() %>% 
  select(taxa, coefficient = coefficient, cilower = cilower, ciupper = ciupper, everything()) %>%
  filter(!str_detect(taxa, "Blastocystis")) %>% # removing metaphlan4's blasto
  filter(!str_detect(taxa, "^GGB")) %>% # removing unclassified
  filter(!str_detect(taxa, "\\d{4}")) %>% # removing unclassified
  # filter(abs(coefficient) > 1) %>%
  ggplot(aes(x = coefficient, y = reorder(taxa, coefficient), group = cohort, color = cohort)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(aes(xmin = cilower, xmax = ciupper, color = cohort), 
                 height = 0.3, alpha = 0.6, position = position_dodge(width = 0.5)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Mean difference in clr-transformed abundance with Blastocystis (RPKSM) abundance", y = "taxa", color = "Cohort") +
  theme_cowplot(14) +
  background_grid()

p_blast
```



### Difference in clr-transformed bacterial abundance ~ Dientamoeba


```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 4

p_dient <- read_csv("data/predict1_dient_interactions.csv") %>% 
  mutate(cohort = "predict1") %>%
  bind_rows(
    read_csv("data/hadza_dient_interactions.csv") %>% 
    mutate(cohort = "hadza"),
    read_csv("data/predict3_uk_dient_interactions.csv") %>% 
    mutate(cohort = "predict3_uk"),
    read_csv("data/predict2_dient_interactions.csv") %>% 
    mutate(cohort = "predict2"),
    read_csv("data/predict3_us1_dient_interactions.csv") %>% 
    mutate(cohort = "predict3_us1"),
    read_csv("data/predict3us22a_dient_interactions.csv") %>% 
    mutate(cohort = "predict3us22a")
  ) %>%
  filter(p.adj.hochberg < p_val_threshold) %>% # only significant taxa
  group_by(taxa) %>%
  filter(n() > cohort_inclusion) %>% # only plotting signicant taxa in at least n cohorts
  ungroup() %>% 
  select(taxa, coefficient = coefficient, cilower = cilower, ciupper = ciupper, everything()) %>%
  filter(!str_detect(taxa, "Dient")) %>% # removing metaphlan4's Dientamoeba
  filter(!str_detect(taxa, "^GGB")) %>% # removing unclassified
  filter(!str_detect(taxa, "\\d{4}")) %>% # removing unclassified
  # filter(abs(coefficient) > 1) %>%
  ggplot(aes(x = coefficient, y = reorder(taxa, coefficient), group = cohort, color = cohort)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(aes(xmin = cilower, xmax = ciupper, color = cohort), 
                 height = 0.3, alpha = 0.6, position = position_dodge(width = 0.5)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Mean difference in clr-transformed abundance with Dientamoeba (RPKSM) abundance", y = "taxa", color = "Cohort") +
  theme_cowplot(14) +
  background_grid()

p_dient
```


### Difference in clr-transformed bacterial abundance ~ Blastocystis adjusting for Dientamoeba and interaction

Effect of Blastocystis RPKSM on bacterial abundance when Dientamoeba is 0.

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 24

p_intxn_blast <- read_csv("data/predict1_itxn_interactions.csv") %>% 
  mutate(cohort = "predict1") %>%
  bind_rows(
    read_csv("data/hadza_itxn_interactions.csv") %>% 
    mutate(cohort = "hadza"),
    read_csv("data/predict3_uk_itxn_interactions.csv") %>% 
    mutate(cohort = "predict3_uk"),
    read_csv("data/predict2_itxn_interactions.csv") %>% 
    mutate(cohort = "predict2"),
    read_csv("data/predict3_us1_itxn_interactions.csv") %>% 
    mutate(cohort = "predict3_us1"),
    read_csv("data/predict3us22a_itxn_interactions.csv") %>% 
    mutate(cohort = "predict3us22a")
  ) %>%
  filter(p.adj.hochberg_blast < p_val_threshold) %>% # only significant taxa
  group_by(taxa) %>%
  filter(n() > 1) %>% # only plotting signicant taxa in at least n cohorts
  ungroup() %>% 
  select(taxa, coefficient = coefficient_blast, cilower = cilower_blast, ciupper = ciupper_blast, everything()) %>%
  filter(!str_detect(taxa, "Blast")) %>% # removing metaphlan4's Blastocystis
  filter(!str_detect(taxa, "Dient")) %>% # removing metaphlan4's Dientamoeba
  filter(!str_detect(taxa, "^GGB")) %>% # removing unclassified
  filter(!str_detect(taxa, "\\d{4}")) %>% # removing unclassified
  # filter(abs(coefficient) > 1) %>%
  ggplot(aes(x = coefficient, y = reorder(taxa, coefficient), group = cohort, color = cohort)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(aes(xmin = cilower, xmax = ciupper, color = cohort), 
                 height = 0.3, alpha = 0.6, position = position_dodge(width = 0.5)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Mean difference in clr-transformed abundance with Blastocystis (RPKSM) abundance when Dientamoeba is absent", y = "taxa", color = "Cohort") +
  theme_cowplot(14) +
  background_grid()

p_intxn_blast
```


### Difference in clr-transformed bacterial abundance ~ Dientamoeba adjusting for Blastocystis and interaction

Effect of Dientamoeba RPKSM on bacterial abundance when Blastocystis is 0.

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 9

p_intxn_dient <- read_csv("data/predict1_itxn_interactions.csv") %>% 
  mutate(cohort = "predict1") %>%
  bind_rows(
    read_csv("data/hadza_itxn_interactions.csv") %>% 
    mutate(cohort = "hadza"),
    read_csv("data/predict3_uk_itxn_interactions.csv") %>% 
    mutate(cohort = "predict3_uk"),
    read_csv("data/predict2_itxn_interactions.csv") %>% 
    mutate(cohort = "predict2"),
    read_csv("data/predict3_us1_itxn_interactions.csv") %>% 
    mutate(cohort = "predict3_us1"),
    read_csv("data/predict3us22a_itxn_interactions.csv") %>% 
    mutate(cohort = "predict3us22a")
  ) %>%
  filter(p.adj.hochberg_dient < p_val_threshold) %>% # only significant taxa
  group_by(taxa) %>%
  filter(n() > 1) %>% # only plotting signicant taxa in at least n cohorts
  ungroup() %>% 
  select(taxa, coefficient = coefficient_dient, cilower = cilower_dient, ciupper = ciupper_dient, everything()) %>%
  filter(!str_detect(taxa, "Blast")) %>% # removing metaphlan4's Blastocystis
  filter(!str_detect(taxa, "Dient")) %>% # removing metaphlan4's Dientamoeba
  filter(!str_detect(taxa, "^GGB")) %>% # removing unclassified
  filter(!str_detect(taxa, "\\d{4}")) %>% # removing unclassified
  # filter(abs(coefficient) > 1) %>%
  ggplot(aes(x = coefficient, y = reorder(taxa, coefficient), group = cohort, color = cohort)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(aes(xmin = cilower, xmax = ciupper, color = cohort), 
                 height = 0.3, alpha = 0.6, position = position_dodge(width = 0.5)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Mean difference in clr-transformed abundance with Dientamoeba (RPKSM) abundance when Blastocystis is absent", y = "taxa", color = "Cohort") +
  theme_cowplot(14) +
  background_grid()

p_intxn_dient

```

### Interaction term in clr-transformed bacterial abundance ~ Blastocystis * Dientamoeba

Blast*Dient coefficient: How much does the Blastocystis effect change depend on Dientamoeba abundance.
If 0, there is no interaction between the two protists. If positive, Dientamoeba amplifies Blastocystis's effect and when negative, Dientamoeba weakens Blastocystis's effect and vice versa.

**No filter of at least in 2 cohorts included in the interaction coefficient figure.**
```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 14

p_itxn_coeff <- read_csv("data/predict1_itxn_interactions.csv") %>% 
  mutate(cohort = "predict1") %>%
  bind_rows(
    read_csv("data/hadza_itxn_interactions.csv") %>% 
    mutate(cohort = "hadza"),
    read_csv("data/predict3_uk_itxn_interactions.csv") %>% 
    mutate(cohort = "predict3_uk"),
    read_csv("data/predict2_itxn_interactions.csv") %>% 
    mutate(cohort = "predict2"),
    read_csv("data/predict3_us1_itxn_interactions.csv") %>% 
    mutate(cohort = "predict3_us1"),
    read_csv("data/predict3us22a_itxn_interactions.csv") %>% 
    mutate(cohort = "predict3us22a")
  ) %>%
  filter(p.adj.hochberg_itxn < 0.01) %>% # only significant taxa
  group_by(taxa) %>%
  # filter(n() > cohort_inclusion) %>% # only plotting signicant taxa in at least n cohorts
  ungroup() %>% 
  select(taxa, coefficient = coefficient_itxn, cilower = cilower_itxn, ciupper = ciupper_itxn, everything()) %>%
  filter(!str_detect(taxa, "Blast")) %>% # removing metaphlan4's Dientamoeba
  filter(!str_detect(taxa, "Dient")) %>% # removing metaphlan4's Dientamoeba
  filter(!str_detect(taxa, "^GGB")) %>% # removing unclassified
  filter(!str_detect(taxa, "\\d{4}")) %>% # removing unclassified
  # filter(abs(coefficient) > 1) %>%
  ggplot(aes(x = coefficient, y = reorder(taxa, coefficient), group = cohort, color = cohort)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(aes(xmin = cilower, xmax = ciupper, color = cohort), 
                 height = 0.3, alpha = 0.6, position = position_dodge(width = 0.5)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Interaction term - Change in Blastocystis effect on clr-transformed abundance with varying Dientamoeba abundance", y = "taxa", color = "Cohort") +
  theme_cowplot(14) +
  background_grid()

p_itxn_coeff
```

### Saving publication quality figures


```{r}
#| warning: false
#| message: false

# code to save high dpi svg/pdfs
save_plot("figures/blast_figure.pdf", p_blast, dpi = 300, base_height = 15, base_width = 12)
save_plot("figures/dient_figure.pdf", p_dient, dpi = 300, base_height = 4, base_width = 12)
save_plot("figures/intxn_blast_figure.pdf", p_intxn_blast, dpi = 300, base_height = 24, base_width = 12)
save_plot("figures/intxn_dient_figure.pdf", p_intxn_dient, dpi = 300, base_height = 9, base_width = 12)
save_plot("figures/itxn_coeff_figure.pdf", p_itxn_coeff, dpi = 300, base_height = 14, base_width = 12)


```